---
title: "Journal"
author: "Roman Neubauer"
date: "2020-11-05"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

**IMPORTANT:** You can delete everything in here and start fresh. You might want to start by not deleting anything above this line until you know what that stuff is doing.

Last compiled: `r Sys.Date()`

This is an `.Rmd` file. It is plain text with special features. Any time you write just like this, it will be compiled to normal text in the website. If you put a \# in front of your text, it will create a top level-header.

# Introduction to the tidyverse
## Challenges
### Analysis of bikes sales by states and cities in Germany
Problem: Analysis of sales by location: display the total revenue of all bikes shops across a state or city in a bar plot.


1. Load needed libraries
```{r, eval = TRUE}
library("tidyverse")
library("readxl")
library("lubridate")
```


2. Importing the data from .xlsx files
```{r, eval = TRUE}
bikes <- read_excel("00_data/01_bike_sales/01_raw_data/bikes.xlsx") %>% set_names(names(.) %>% str_replace_all("\\.", "_")) %>% select(bike_id, model, price)
shops <- read_excel("00_data/01_bike_sales/01_raw_data/bikeshops.xlsx") %>% set_names(names(.) %>% str_replace_all("\\.", "_"))
orders <- read_excel("00_data/01_bike_sales/01_raw_data/orderlines.xlsx") %>% select(- ...1) %>% set_names(names(.) %>% str_replace_all("\\.", "_"))
```
in the same step some unneeded columns were dropped and all dots in column names were replaced by underscores.


3. Joining the data sets and wrangling
```{r, eval = TRUE}
orders_shops_joined <- left_join(orders, shops, by=c("customer_id" = "bikeshop_id")) %>% select(-lat, -lng, -order_line)

joined_data <- left_join(orders_shops_joined, bikes, by=c("product_id"="bike_id")) %>% 
  #separating city and state
  separate(col=location, into=c("city", "state"), sep=", ") %>% 
  #calculating total order price
  mutate(total_price = price*quantity) %>% 
  #adding year of the sale
  mutate(year = year(order_date))
```


4. Manipulating

getting the sales per state:
```{r, eval = TRUE}
sales_by_state <- joined_data %>% 
  #select needed columns
  select(state, total_price) %>% 
  #grouping by state
  group_by(state) %>% 
  #adding all revenues for the given states
  summarize(revenue = sum(total_price)) %>% 
  #formatting the revenue 
  mutate(revenue_text = scales::dollar(revenue, big.mark = ".", decimal.mark = ",", prefix = "", suffix = " €"))
```

getting the sales by city
````{r, eval=TRUE}
sales_by_city <- joined_data %>% 
  #select needed columns
  select(city, total_price) %>% 
  #grouping by city
  group_by(city) %>% 
  #adding all revenues for the given city
  summarize(revenue = sum(total_price)) %>% 
  #formatting the revenue
  mutate(revenue_text = scales::dollar(revenue, big.mark = ".", decimal.mark = ",", prefix = "", suffix = " €"))
````

5. Plotting the data

for the sales by state:
```{r, eval=TRUE}

sales_by_state %>%
  
  # Setup canvas with the columns states (x-axis) and revenue (y-axis)
  ggplot(aes(x = state, y = revenue)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") +
  geom_label(aes(label = revenue_text), size=2) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  
  labs(
    title    = "Revenue by state over 5 years",
    x = "State",
    y = "Revenue"
  )
```

for the sales by city:
```{r, eval=TRUE}
sales_by_city %>%
  
  # Setup canvas with the columns states (x-axis) and revenue (y-axis)
  ggplot(aes(x = city, y = revenue)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") +
  geom_label(aes(label = revenue_text), size=2) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  
  labs(
    title    = "Revenue by city over 5 years",
    x = "City",
    y = "Revenue"
  )
```

### Analysis of bike sales by state and year
Steps 1 to 3 remain the same as in 1. 

4. Manipulating
getting the sales by state and year:
```{r, eval=TRUE}
sales_by_state_and_year <- joined_data %>% 
  #select needed columns
  select(state, year, total_price) %>% 
  #grouping by state and year
  group_by(state, year) %>% 
  #adding all revenues for a given state and year
  summarize(revenue = sum(total_price)) %>% 
  #formatting the revenue
  mutate(revenue_text = scales::dollar(revenue, big.mark = ".", decimal.mark = ",", prefix = "", suffix = " €"))
```

5. Plotting
```{r plot, fig.width=12, fig.height=7}
sales_by_state_and_year %>%
  
  # Set up x, y, fill
  ggplot(aes(x = year, y = revenue, fill = state)) +
  
  # Geometries
  geom_col() +
  
  # Facet
  facet_wrap(~ state) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  
  labs(
    title = "Revenue by state and year",
    subtitle = "displayed revenue for all  bikeshops in a given state",
    fill = "States"
  )
```



# Data Acquisition



# Examples
Notice that whatever you define as a top level header, automatically gets put into the table of contents bar on the left. 

Adding this code to my first post:
```{r, eval = FALSE}
numbers <- 1:1000

# This will print the first 10 elements of the vector numbers
numbers[1:10]

# This will plot a histogram of 100 random elements of the vector numbers
hist(sample(numbers, 100, replace = T))
```

## Second level header

You can add more headers by adding more hashtags. These won't be put into the table of contents

### third level header

Here's an even lower level header

# My second post (note the order)

Last compiled: `r Sys.Date()`

I'm writing this tutorial going from the top down. And, this is how it will be printed. So, notice the second post is second in the list. If you want your most recent post to be at the top, then make a new post starting at the top. If you want the oldest first, do, then keep adding to the bottom

# Adding R stuff

So far this is just a blog where you can write in plain text and serve your writing to a webpage. One of the main purposes of this lab journal is to record your progress learning R. The reason I am asking you to use this process is because you can both make a website, and a lab journal, and learn R all in R-studio. This makes everything really convenient and in the same place. 

So, let's say you are learning how to make a histogram in R. For example, maybe you want to sample 100 numbers from a normal distribution with mean = 0, and standard deviation = 1, and then you want to plot a histogram. You can do this right here by using an r code block, like this:

```{r}
samples <- rnorm(100, mean=0, sd=1)
hist(samples)
```

When you knit this R Markdown document, you will see that the histogram is printed to the page, along with the R code. This document can be set up to hide the R code in the webpage, just delete the comment (hashtag) from the cold folding option in the yaml header up top. For purposes of letting yourself see the code, and me see the code, best to keep it the way that it is. You'll learn that all of these things and more can be customized in each R code block.